---
title: "Obligatorisk 6"
author: "Adrian"
date: "1 november 2018"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Loading all the required packages

```{r}
library(tidyr)
library(tidyverse)
library(dplyr)
```
Task 1: Read the following accounting dataset DS: 
```{r}
DS = tribble(
 ~fk_account_code, ~Ansvar, ~fk_function_code, ~fk_project_code, ~amount,
 101030,40220,"Det",3432,1493.00,
 101030,40220,"Met",3586,2827.00,
 101030,40320,"Det",3456,49440.00,
 101030,40330,"Sal",NA,870716.00,
 101030,40350,"Met",NA,559928.00,
 101030,40360,"Sal",NA,125534.00,
 101030,40280,"Pol",NA,251611.00)

```
1) Remove the "fk_project_code" variable from DS.
```{r}
glimpse(DS)
DS["fk_project_code"]<- NULL
```
2) Sum the "amount" by the 3 first digits of "Ansvar"; 402, 403, etc.
```{r}
DS$Ansvar <- as.character(DS$Ansvar)
str(DS)

DS <- separate(DS, Ansvar, into = c("Ansvar 1_3", "rest"), sep = 3)
DS %>% group_by(`Ansvar 1_3`) %>% summarise(sum=sum(amount))
str(DS)
```
3) Make new "labels" for "fk_function_code", where:
 # "Det" and "Sal" is "supplies",
 #"Met" is "inventories" and
 #"Pol" is "other expenses"
```{r}
DS %>% mutate(fk_function_code = case_when(fk_function_code== "Det"~ "supplies",
                                           fk_function_code== "Sal" ~ "supplies",
                                           fk_function_code== "Met" ~ "inventories",
                                           fk_function_code== "Pol" ~ "other expenses"))
```
 
TASK 2: 
Read the following small dataset df:
```{r}

df <- data.frame(Product=gl(3,10,labels=c("A","B", "C")), 
                   Year=factor(rep(2002:2011,3)), 
                   Sales=1:30)
```
#Calculate the share of sales per product per year. The sum over the 3 shares per year is 100.
#Make a plot of the sales and shares per year per company.
```{r}
library(data.table)
DT <- data.table(df)
str(DT)
#Using data.table, choose an index variable, key variable
setkey(DT, "Year")
#calcualte shares by year
DT<- DT[, share := Sales/(sum(Sales)), by=list(Year)]
View(DT)

library(latticeExtra)
xyplot(Sales +share ~ Year, groups=Product, data = DT)

```