---
title: "ob_3"
author: "Adrian"
date: "7 oktober 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("PxWebApiData", repos = "https://cran.r-project.org/package=PxWebApiData")
install.packages("tidyquant", repos = "https://cran.r-project.org/package=tidyquant")
library(readr)
library(knitr)
library(tidyquant)
library(ggplot2)
library(tidyverse)
library(tidyr)
library(dplyr)
library(PxWebApiData)
options(encoding="UTF-8")
```
Getting data from every county
```{r}
hvert_fylke <- ApiData("http://data.ssb.no/api/v0/dataset/95274.json?lang=no", getDataByGET = TRUE, col_types = cols(`måned` = col_date(format = "%Y%m")))
```
Doing some changes to make it easier to read: 
```{r}
hvert_fylke <- hvert_fylke[[1]]
hvert_fylke<-subset(hvert_fylke, value>0)
names(hvert_fylke) <-c("region", "date", "statistikkvariabel", "value")
hvert_fylke
```
```{r}
hvert_fylke <- separate(hvert_fylke, date, c("year", "month"), sep = "M")
hvert_fylke <- unite(hvert_fylke, date, c("year", "month"), sep = "-")
```
From "long" to "wide"
```{r}
hvert_fylke <- spread(hvert_fylke, statistikkvariabel, value)
```
Fixing the date format
```{r}
hvert_fylke <- mutate(hvert_fylke, date= as.Date(date, format= "%Y"))
nytid <- seq(as.Date("1985-01-01"), as.Date("2018-07-01"), by = "months")
repnytid <- rep(nytid, length.out = 7750)
hvert_fylke$date <- repnytid
```
Removing unnecessary variables. 
```{r}
names(hvert_fylke) <-c("region", "date", "rom", "senger", "pris")
hvert_fylke$rom <- NULL 
hvert_fylke$senger <- NULL
hvert_fylke<-subset(hvert_fylke, pris > 0)
names(hvert_fylke) <-c("region", "date", "fylkepris")
```
Making a ggplot from 1985 to 2018
```{r}
min <- as.Date("1985-01-01")
max <- as.Date("2018-07-01")


ggplot2::ggplot(hvert_fylke, aes(x = fylkepris, y = region)) +
  geom_line(color = "#00AFBB")
```

Svalberg is the most expensive out of all the counties. 


#Doing the same thing with the entire country. 
Getting data: 
```{r}
land <- ApiData("http://data.ssb.no/api/v0/dataset/95276.json?lang=no", getDataByGET = TRUE)
```
Making "month" to "date"
```{r}
hele_landet <- land[[1]]
hele_landet<-subset(hele_landet, value>0)
names(hele_landet) <-c("region", "date", "statistikkvariabel", "value")
hele_landet
```
Seperate M, like I did with county. 
```{r}
hele_landet <- separate(hele_landet, date, c("year", "month"), sep = "M")
hele_landet <- unite(hele_landet, date, c("year", "month"), sep = "-")
```

Changes from "long" to "wide" here too
```{r}
hele_landet <- spread(hele_landet, statistikkvariabel, value)
```
using nytid to fix date
```{r}
hele_landet <- mutate(hele_landet, date= as.Date(date, format= "%Y"))
nytid <- seq(as.Date("1985-01-01"), as.Date("2018-07-01"), by = "months")
repnytid <- rep(nytid, length.out = 404)
hele_landet$date <- repnytid
```

make sure I only have "pris per rom" and "dato", removing all the 0 values
```{r}
names(hele_landet) <-c("region", "date", "rom", "senger", "pris")
hele_landet$region <- NULL
hele_landet$rom <- NULL
hele_landet$senger <- NULL
hele_landet<-subset(hele_landet, pris > 0)
names(hele_landet) <-c("date", "prisnorge")
```
Looking over the data using "head", "glimpse" and "tail".
```{r}
head(hele_landet, 35)
tail(hele_landet, 35)
glimpse(hele_landet)
```
now to make a ggplot for "hele_landet"
```{r}
ggplot2::ggplot(hele_landet, aes(x = date, y = prisnorge)) +
  geom_line(color = "#E7B800")
```


Now to merge both datas and see if there is any correlation between "hele_landet" and "hvert_fylke"
```{r}
hvert_fylke$region <- NULL
merge <- left_join(hvert_fylke, hele_landet, by=c("date"))
```

```{r}
merge.long <- merge %>% gather(key = "variable", value = "pris", -date)
head(merge.long)
```








making a ggplot for the entire country. 
```{r}
ggplot(merge.long, aes(x = variable, y = pris)) + 
  geom_line(aes(color = variable)) +
  scale_color_manual(values = c("#00AFBB", "#E7B800"))
```
```{r}
ggplot(merge.long, aes(x = date, y = pris)) + 
  geom_line(aes(color = variable)) +
  scale_color_manual(values = c("#00AFBB", "#E7B800"))
```













